Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-05-31 supp_mat_complexity_selection.Rmd
updating figures with relative fitness and bootstrapped confidence intervals.

## CREATE SEPARATE DATASETS FOR SELECTION ANALYSES ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") %>%
  mutate(Gall_Height_mm = Gall_Height_mm - mean(Gall_Height_mm),
         gall_individuals = gall_individuals - mean(gall_individuals),
         Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") %>%
  mutate(Gall_Height_mm = Gall_Height_mm - mean(Gall_Height_mm),
         gall_individuals = gall_individuals - mean(gall_individuals),
         Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))
```

@@

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

@@

>>>>

## CREATE SEPARATE DATASETS FOR SELECTION ANALYSES ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))
```

@@

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

inverse_logit <- function(x) exp(x)/(1+exp(x))
```

@@

## Control treatment
Wbar_model_control <-  glmer(gall_survival ~ 1 + (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = control_df, glmerControl(optimizer="bobyqa"))
Wbar_control <- inverse_logit(fixef(Wbar_model_control))

glmer_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = control_df, glmerControl(optimizer="bobyqa"))
summary(glmer_control)
```

## Ectoparasitoid exclusion treatment
Wbar_model_treatment <-  glmer(gall_survival ~ 1 + (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = treatment_df, glmerControl(optimizer="bobyqa"))
Wbar_treatment <- inverse_logit(fixef(Wbar_model_treatment))

glmer_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference + 
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = treatment_df, glmerControl(optimizer="bobyqa"))
summary(glmer_treatment)
```

```{r}
distribution_female_preference <- distinct(.data=gall_selection.df, Plant_Position, Density_per_100_shoots)$Density_per_100_shoots 
HPDinterval_female_preference <- coda::HPDinterval(coda::as.mcmc(scale(distribution_female_preference)), prob=0.95); HPDinterval_female_preference

bind_rows(mutate(visreg(glmer_treatment, xvar="sc.female_preference", scale="response", plot=F)$fit, 
                   Treatment="Simple", `Relative fitness` = visregFit/Wbar_treatment), 
          mutate(visreg(glmer_control, xvar="sc.female_preference", scale="response", plot=F)$fit, 
                   Treatment="Complex", `Relative fitness` = visregFit/Wbar_control)) %>%
  ggplot(., aes(x=sc.female_preference, y=`Relative fitness`, color=Treatment)) +
  geom_line() +
  scale_x_continuous(limits = c(HPDinterval_female_preference[1], HPDinterval_female_preference[2]))
```

```{r}
distribution_clutch_size <- distinct(.data=gall_selection.df, Plant_Position, Gall_Number, gall_individuals)$gall_individuals
HPDinterval_clutch_size <- coda::HPDinterval(coda::as.mcmc(scale(distribution_clutch_size)), prob=0.95); HPDinterval_clutch_size

bind_rows(mutate(visreg(glmer_treatment, xvar="sc.clutch_size", scale="response", plot=F)$fit, 
                   Treatment="Simple", `Relative fitness` = visregFit/Wbar_treatment), 
          mutate(visreg(glmer_control, xvar="sc.clutch_size", scale="response", plot=F)$fit, 
                   Treatment="Complex", `Relative fitness` = visregFit/Wbar_control)) %>%
  ggplot(., aes(x=sc.clutch_size, y=`Relative fitness`, color=Treatment)) +
  geom_line() +
  scale_x_continuous(limits = c(HPDinterval_clutch_size[1], HPDinterval_clutch_size[2]))
```

```{r}
distribution_gall_size <- distinct(.data=gall_selection.df, Plant_Position, Gall_Number, Gall_Letter, Gall_Height_mm)$Gall_Height_mm
HPDinterval_gall_size <- coda::HPDinterval(coda::as.mcmc(scale(distribution_gall_size)), prob=0.95); HPDinterval_gall_size

predict_gall_size_df <- data.frame(sc.gall_size = seq(HPDinterval_gall_size[1], HPDinterval_gall_size[2], length.out=100),
                                   sc.clutch_size = 0,
                                   sc.female_preference = 0)
model_matrix_gall_size <- model.matrix(~sc.gall_size + sc.clutch_size + sc.female_preference, predict_gall_size_df)

# ref: https://www.r-bloggers.com/confidence-intervals-for-prediction-in-glmms/
boot_treatment_gall_size <-bootMer(glmer_treatment, FUN = function(.) inverse_logit(model_matrix_gall_size %*% fixef(.)), nsim=200)
se_treatment_gall_size <- apply(boot_treatment_gall_size$t, 2, function(x) x[order(x)])
boot_lower <- se_treatment_gall_size[1, ]
boot_upper <- se_treatment_gall_size[2, ]

bind_rows(mutate(visreg(glmer_treatment, xvar="sc.gall_size", scale="response", plot=F)$fit, 
                   Treatment="Simple", `Relative fitness` = visregFit/Wbar_treatment), 
          mutate(visreg(glmer_control, xvar="sc.gall_size", scale="response", plot=F)$fit, 
                   Treatment="Complex", `Relative fitness` = visregFit/Wbar_control)) %>%
  ggplot(., aes(x=sc.gall_size, y=`Relative fitness`, color=Treatment)) +
  geom_line() +
  geom_ribbon(aes(ymin=boot_lower, ymax=boot_upper)) +
  scale_x_continuous(limits = c(HPDinterval_gall_size[1], HPDinterval_gall_size[2]))
```
